\section{General Rules}


\subsection{Procedure Manuals and Officials}\label{manuals}

This manual governs procedure at all \FGCCFL-sponsored tournaments. 
Consult this manual first, 
followed by the most recent \emph{NCFL Student Congress Manual}, 
then \emph{Robert's Rules of Order, Newly Revised} (11th ed.).

Judges (Parliamentarians at Grand Finals) may make rulings at the request of the \po\ 
but otherwise should only intervene if necessary to correct significant mistakes. 
Decisions made in the chamber \willnot\ be overturned except for clear error.

The Congress Director and tab staff may make rulings at the request of the judges 
and may take all actions necessary to ensure a fair and efficient tournament.


\subsection{Conduct and Decorum}\label{conduct}

\textit{\textbf{Student Congress chambers should embody the best ideals of 
representative democracy and deliberative policymaking. 
Participants are expected to treat the officials and one another 
with dignity and respect at all times.} 
Personal attacks, harassment, or bullying of any sort may result in sanctions 
up to and including disqualification from the tournament.}

\begin{enumerate}

\item
Legislators must remain silent when others have the floor, 
and the \po\ (PO) must rule interruptions \ooo\ except when 
(i) permitted by the rules of procedure 
\andd\ (ii) in furtherance of the chamber's needs.
Legislators \may\ pass notes and share materials \emph{discreetly and quietly}.

\item
Speakers shall direct their remarks to the chamber as a whole, 
not to individual members. 
Legislators are to address and refer to each other as 
``Representative \emph{(surname)}'' or ``Mr./Ms. \emph{(surname)}.'' 
The PO is addressed as ``Madam/Mister Speaker'' and is referred to as ``the Chair.''

\item
Legislators \must\ attend \all\ sessions in their assigned chambers, 
except for \emph{very brief} exits from the chamber when absolutely necessary. 
The PO shall deny excessive requests to exit the chamber, and the chamber \maynot\ 
suspend the rules to permit ``open'' exit and re-entry.

\item
Legislators are held to the highest ethical standards. 
Falsification of evidence may result in penalties up to forfeiture of the session. 
Judges may review evidence in response to an allegation of falsification if 
the original source is available in the chamber.

\end{enumerate}


\subsection{The Presiding Officer (PO)}\label{the_po}
The PO's job is to lead the chamber in productive and fair legislative discussion. 
To that end, she\footnote{ 
	This manual uses feminine pronouns when referring to Presiding Officers and 
	masculine pronouns when referring to other members. All pronouns should be 
	considered gender-inclusive.
} 
\textbf{may and should} advise the chamber on matters of procedure 
not related to the substance of any item of legislation. 
She should also be careful to recognize speakers fairly 
and to ensure that legislators maintain decorum.

Furthermore, the PO is strictly charged with the duty to keep debate moving 
and has substantial discretion to do so.
She \must\ rule dilatory tactics and frivolous motions \ooo\ 
and \may\ deny recognition to legislators attempting to monopolize the floor. 
These rulings are subject to appeal.

The PO \mustnot\ enter into debate on legislation, 
and she \maynot\ relinquish the Chair to participate in floor debate. 
Legislators may not ``approach the Chair'' or pass notes to the PO, 
except to submit an amendment for review.

The PO should stand when addressing the chamber if she is able to do so.


\subsection{Electronic Devices}\label{laptops}


























